# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data/burden-of-disease-cmnn.csv"))
```

```{r}
# Filter the dataset to include only the three countries chosen by your group.
data_cmnn_mini <- data_cmnn %>% 
                  filter( Entity %in% c("Angola", 
                        "China", 
                        "Benin") ) %>% 
  rename(daly_burden = `DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`)
  
```

## Table of Estimates for CMNN Burden Over Time

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Here render a table for the DALY burden over time for the three countries.
# You should pivot the data wider to show each country in a separate column.
data_cmnn_wide <- data_cmnn_mini %>% 
  select(-Code) %>% 
  pivot_wider(names_from = Entity, 
              values_from = daly_burden) %>% 
# Use kable() from the knitr package to render the table.

kable()

data_cmnn_wide
```

## Plot Showing Trends in CMNN Burden Over Time

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
ggplot(data = data_cmnn_mini, 
       aes(x = Year, 
           y = daly_burden, 
           
           colour = Entity,
           group = Entity
           )) + 
  geom_line(size = 1) + 
  labs(
    title = "DALY Burden Over Years by Country",
    x = "Year",
    y = "DALY Burden",
    color = "Country")
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table.
```

## Summary of CMNN Burden Findings

Globaly, The plot shows a clear decline in DALY burden from 1990 to 2019, indicating improvements in public health and disease control. Angola had the highest burden initially but experienced the steepest reduction, while Benin declined more gradually, and China maintained the lowest burden throughout.

## Attribution

This section was completed by **Bienvenu Kabasele**

The link to the group's GitHub repository is *<https://github.com/bkabaseleilunga-bit/rmc_week_02_workshop>*
