# Non-Communicable Disease Burden (NCD)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data/burden-of-disease-ncd.csv"))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Filter the dataset to include only the three countries chosen by your group.
library(janitor)
data_ncd_mini <- data_ncd %>% filter(Code=="AGO" | Code=="CHN" | Code=="BEN") 
  colnames(data_ncd_mini)[4] <- "burden"
  
```

## Table of Estimates for NCD Burden Over Time

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Here render a table for the DALY burden over time for the three countries.
# You should pivot the data wider to show each country in a separate column.
data_ncd_wide <-  data_ncd_mini %>% 
  select(-Code) %>% 
  pivot_wider(
    names_from  = Entity,
    values_from = burden
  ) %>% 
# Use kable() from the knitr package to render the table.
kable()
data_ncd_wide

```

## Plot Showing Trends in NCD Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table.
ggplot(data=data_ncd_mini,
       mapping = aes(x = Year, y= burden, color=Entity))+
  geom_line()
```

## Summary of NCD Burden Findings

We observed that all three countries had a high burden of non-communicable diseases in the 1990s. However, over the years, we observed a decrease in the burden of non-communicable diseases in all these countries. That being said, China experienced a more pronounced decline compared with Angola and Benin.

## Attribution

This section was completed by **Fara**

The link to the group's GitHub repository is *<https://github.com/bkabaseleilunga-bit/rmc_week_02_workshop>*
